---
title: "Using priorsense with JAGS"
vignette: >
  %\VignetteIndexEntry{Using priorsense with JAGS}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---


```{r}
#| include: false
ggplot2::theme_set(bayesplot::theme_default(base_family = "sans"))

options(priorsense.plot_help_text = FALSE)
```


```{r}
#| message: false
#| warning: false
library(nimble)
library(posterior)
library(priorsense)
```

To use `priorsense` with a NIMBLE model, the log prior and log likelihood
evaluations should be added to the model code.

```{r}
code <- nimbleCode({
  for (n in 1:N) {
    y[n] ~ dnorm(mu, sd = sigma)
    log_lik[n] <- dnorm(y[n], mu, sd = sigma, log = TRUE)
  }

  mu ~ dnorm(0, sd = 1)
  sigma ~ dnorm(0, sd = 2.5)

  lprior_mu <- dnorm(mu, 0, sd = 1, log = TRUE)
  lprior_sigma <- dnorm(sigma, 0, sd = 2.5, log = TRUE)

  lprior <- lprior_mu + lprior_sigma

})
```
Instantiate and compile the model as usual.
```{r}
#| message: false
#| warning: false

inits <- list(
  mu = 0,
  sigma = 1
)

model <- nimbleModel(
  code,
  data      = d,
  inits     = inits,
  constants = list(N = d$N)
)


cmodel <- compileNimble(model)

mcmc <- buildMCMC(
  model,
  monitors = c(
    "mu",
    "sigma",
    "lprior",
    "lprior_mu",
    "lprior_sigma",
    "log_lik"
  )
)

cmcmc <- compileNimble(mcmc, project = model)
```

For priorsense compatibility, it is easiest to save the draws as a `coda::mcmc.list` object, by specifying `samplesAsCodaMCMC = TRUE`. Otherwise, the resulting object can be transformed to a `posterior::draws_df` object, with `posterior::as_draws_df`.

```
#| message: false
#| warning: false

nimble_fit <- runMCMC(
  cmcmc,
  niter = 20000,
  nburnin = 5000,
  nchains = 4,
  thin = 5,
  setSeed = c(123, 456, 789, 101112),
  samplesAsCodaMCMC = TRUE # alternatively, coerce the output using `posterior::as_draws_df`
)

```

Then the `priorsense` functions will work as usual.

```{r}
powerscale_sensitivity(nimble_fit)
```

```{r}
#| message: false
#| warning: false
#| fig-width: 6
#| fig-height: 4
powerscale_plot_dens(nimble_fit)
```
